---
title: "dplyr"
author: "Cameron Nemo"
date: "February 13, 2018"
output: html_document
---

```{r setup, message=FALSE}
library(dplyr)
library(hflights)
```

## data structures: tibbles and lookup tables

Tibbles are an efficient and convenient way to model data. The `as_tibble` function can be used to transform a data frame, matrix, or list to a tbl. A tbl can be operated on just like a data frame. 

```{r}
# Convert the hflights df to a tbl, tfly
tfly <- as_tibble(hflights)
```

Lookup tables are useful in transforming a variable

```{r}
# Create carrier codes from a lookup table
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental",
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")
unique(tfly$UniqueCarrier)
unique(lut[tfly$UniqueCarrier])
```

## five verbs

These functions manipulate the contents of the data, but do not modify in place. Their results must be stored to be persistent.

When using column names, there is no need to prefix the tibble name. The verbs process their arguments in the context of the dataframe.

### filter

Subset by row. Only observations matching the logical expression are included.

```{r}
# Get average delay of each carrier in 2011
tfly %>% filter(Year == 2011) %>% group_by(UniqueCarrier) %>% summarise(AvgDepDelay = mean(DepDelay, na.rm=TRUE))
```

### arrange

Orders the observations according to (a) given variable(s). Priority for first stated variables is higher.

`desc()` useful for descending order.

```{r}
# Group by carrier, and within each carrier by departure delay
tfly %>% arrange(UniqueCarrier, DepDelay)
# Order by descending departure delay
tfly %>% arrange(desc(DepDelay))
```

### select

Subset by column, can use ranges and can negate columns with `:` and `-`, respectively.

```{r}
# Available column names (variables)
tfly %>% colnames
# Subset the variables of the tbl
tfly %>% select(Year:ArrTime, -DayofMonth) %>% colnames
```

### mutate

Create new columns using existing information in a dataset. Argument(s) are `var_name=expression`.

```{r}
# Create a summed delay time variable
tfly <- tfly %>% mutate(TotalDelay = ArrDelay + DepDelay)
tfly %>% select(ends_with("Delay"))
# Rank the delay times of each flight from shortest to longest
tfly %>% mutate(DelayRank = rank(TotalDelay)) %>% select(contains("Delay"))
```

### summarise

Creates a table of summary statistics for the tbl or each group. Unlike mutate, collapses existing information.

```{r}
# Mean and variance for all the flights in the tbl
tfly %>% summarise(avg = mean(TotalDelay, na.rm=TRUE), var = var(TotalDelay, na.rm=TRUE))
# Number of flights grouped by carrier
tfly %>% group_by(UniqueCarrier) %>% summarise(flight_count = n())
```

## dplyr helper functions

Useful for select: 

* `starts_with(s)`, `ends_with(s)`, `contains(s)`, `matches(s)`, where s is a string to match
* `num_range(prefix, range)`, where prefix is a string and range is vector of numbers
* `one_of(vector)`, where vector is a string vector

Useful for summarise:
 
* `first(x)`, `last(x)`, `nth(x, n)` (element of a vector, x)
* `n()`: number of observations in the tbl
* `n_distinct(x)`: number of unique observations in the vector x, same as `length(unique(x))`
* `group_by(...)`: takes one or more arguments to form groups

`tally()`: wrapper for summarise, provides number of observations

`count(...)`: like tally, but automatically groups and ungroups

## notable base functions

### rank

```{r}
vec <- seq(1, 10, by=2)
rank(vec)
rank(desc(vec))
```

## database operations

