---
title: "Data Visualisation with ggplot2"
author: "Cameron Nemo"
date: "February 21, 2018"
output: html_document
---

# 0. Introduction

```{r message=FALSE}
library(dplyr)
library(ggplot2)
```

`ggplot2` uses seven grammar elements form the layers of its visualisations.

# 1. Data

The first argument to `ggplot()` is the data set being visualized.

```
mt_plot <- mtcars %>% ggplot()
```

# 2. Aesthetics, and attributes

Next, use an aesthetic mapping to link variables with visual elements in the plot.

```{r}
mt_plt <- mtcars %>% ggplot(aes(x=disp, y=mpg))
```

Aesthetics differ from attributes in that they are related to variables in the data, while attributes are constant.

Aesthetics available include:

* `x` and `y` axes
* `color`, `fill`, `shape`, `linetype` usually used for categorical variables, but can be applied to continuous variables
* `size` and `alpha` control size and transparency of shapes in the plot, usually used for continuous variables
* `alpha` controls transparency of data elements in a plot
* labels are also available

While `color` applies to the outline of shapes, `fill` applies to the interior of shapes.

On the `shape` aesthetic: it can have a value from 1-25. 1-20 only accept a `color` aesthetic, but 21-25 have both a `color` and a `fill` aesthetic. For example, when working with `geom_point`, setting shape to 21 allows one to use both the col and fill aesthetics.

```{r}
ggplot(mtcars, aes(x=wt, y=mpg, fill=cyl, col=am)) +
  geom_point(shape=21, size=4, alpha=0.6)
```

## Modifying aesthetics

position and scale functions can be used to modify aesthetics.

position is usually set within the geom's function, but can also be added as a separate layer. For a bar chart, the default position is `"stack"`, but other options include `"fill"` for a proportional bar chart and `"dodge"` for side by side bars.

scale functions are in the format `scale_AESTHETIC_METHOD`, such as `scale_x_continuous` or `scale_fill_manual`.

The first argument, `scale_name` is the name of the scale that is used to label the axis or legend. By default it is the name of the variable mapped to the aesthetic. Other arguments include `values`, `labels`, `limits`, and `breaks`.

# 3. Geometries

# 4. Facets

# 5. Statistics

# 6. Coordinates

# 7. Themes

All non-data ink.